Generalized Algebraic Data Types in Haskell

### 2013-06-04 gadt.tex
Minor tweaks and two new comments.
Signed-off-by: Edward Z. Yang <ezyang@mit.edu>

On the one hand, the count parameter must be passed as a value to populate the list at run-time. On the other hand, we need a type-level representation of the same number for \verb|List| type. Haskell enforces a phase separation between run-time values and compile-time types. This causes a problem with type signature of function \verb|repeatElem|: we can not specify type-level representation of list elements count in the return value type of the function matching the count passed as a value.

@@

We have shown that insertion operation maintains all invariants in red-black trees by adding necessary properties to the tree data type and then modifying the functions accordingly. However, the previous code illustrates a general problem with proofs. In fact, in Haskell $\bot$ (bottom) is a member of every type. As a result, we can write:

@@

Type inference for programs with GADTs is out of scope for this article. But when writing programs we are faced with the problem: typically Haskell functions do not require writing type signatures, but many functions involving GADTs do not compile without specifying type signatures explicitly, because Haskell compiler can not automatically infer the type. This section describes the reasons of this situation.

@@

>>>>

On the one hand, the count parameter must be passed as a value to populate the list at run-time. On the other hand, we need a type-level representation of the same number for \verb|List| type. Haskell enforces a phase separation between run-time values and compile-time types. This causes a problem with the type signature of the function \verb|repeatElem|: there is no direct way to specify a type-level representation of list's count in the return value type of the function which will match the count passed in as a value.

@@

\EZY{The new sentence you added here is really awkward} We have shown that the insertion operation maintains all invariants in red-black trees by adding necessary properties to the tree data type and then modifying the functions accordingly. However, the previous code illustrates a general problem with proofs. In fact, in Haskell $\bot$ (bottom) is a member of every type. As a result, we can write:

@@

Type inference for programs with GADTs is out of scope for this article. But when writing programs we are faced with a problem: typically Haskell functions do not require writing type signatures, but many functions involving GADTs do not compile without specifying type signatures explicitly, as the Haskell compiler can not automatically infer the type. This section describes the reasons of this situation.

@@

\EZY{You should probably cite Data Types a la Carte here}
